---
title: "Geocoded Individual Data Joined With Environmental Data"
author: "Daniel Antal, CFA"
date: "3/5/2021"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-pkgs, message=FALSE, warning=FALSE}
library(regions)
library(lubridate)
library(dplyr)

if ( dir.exists('data-raw') ) {
  data_raw_dir <- "data-raw"
} else {
  data_raw_dir <- file.path("..", "..", "data-raw")
  }
```

The first results of our longitudinal table [were difficult to map](post/2021-03-05-retroharmonize-climate/), because the surveys used an obsolete regional coding. We will adjust the wrong coding, when possible, and join the data with the European Environment Agency's (EEA) Air Quality e-Reporting (AQ e-Reporting) data on environmental pollution. We recoded the  annual level for every available reporting stations [*not shown here*] and all values are in μg/m3. The period under observation is 2014-2016. Data file: [https://www.eea.europa.eu/data-and-maps/data/aqereporting-8](https://www.eea.europa.eu/data-and-maps/data/aqereporting-8) (European Environment Agency 2021).

## Recoding the Regions

Recoding means that the boundaries are unchanged, but the country changed the names and codes of regions because there were other boundary changes which did not affect our observation unit. We explain the problem and the solution in greater detail in [our tutorial](http://netzero.dataobservatory.eu/post/2021-03-06-regions-climate/) that aggregates the data on regional levels.

```{r recoding}
panel <- readRDS((file.path(data_raw_dir, "climate-panel.rds")))

climate_data_geocode <-  panel %>%
  mutate ( year = lubridate::year(date_of_interview)) %>%
  recode_nuts()
```

Let's join the air pollution data and join it by corrected geocodes:

```{r joining}
load(file.path("data", "air_pollutants.rda")) ## good practice to use system-independent file.path

climate_awareness_air <- climate_data_geocode %>%
  rename ( region_nuts_codes  = .data$code_2016) %>%
  left_join ( air_pollutants, by = "region_nuts_codes" ) %>%
  select ( -all_of(c("w1", "wex", "date_of_interview", 
                     "typology", "typology_change", "geo", "region"))) %>%
  mutate (
    # remove special labels and create NA_numeric_ 
    age_education = retroharmonize::as_numeric(age_education)) %>%
  mutate_if ( is.character, as.factor) %>%
  mutate ( 
    # we only have responses from 4 years, and this should be treated as a categorical variable
    year = as.factor(year) 
    ) %>%
  filter ( complete.cases(.) ) 
```

The `climate_awareness_air` data frame contains the answers of `r nrow(climate_awareness_air)` individual respondents. `r paste0(round(mean(climate_awareness_air$serious_world_problems_first)*100,2), "%")` thought that climate change was the most serious world problem and `r paste0(round(mean(climate_awareness_air$serious_world_problems_climate_change)*100,2), "%")` mentioned climate change as one of the three most important global problems.

```{r summary-dataset}
summary ( climate_awareness_air  )
```
Let's see a simple CART tree! We remove the regional codes, because there are very serious differences among regional climate awareness. These differences, together with education level, and the year we are talking about, are the most important predictors of thinking about climate change as the most important global problem in Europe.

```{r rpart}
# Classification Tree with rpart
library(rpart)

# grow tree
fit <- rpart(as.factor(serious_world_problems_first) ~ .,
   method="class", data=climate_awareness_air %>%
     select ( - all_of(c("rowid", "region_nuts_codes"))), 
   control = rpart.control(cp = 0.005))

printcp(fit) # display the results
plotcp(fit) # visualize cross-validation results
summary(fit) # detailed summary of splits

# plot tree
plot(fit, uniform=TRUE,
   main="Classification Tree: Climate Change Is The Most Serious Threat")
text(fit, use.n=TRUE, all=TRUE, cex=.8)
```

```{r save-to-tempdir}
saveRDS ( climate_awareness_air , file.path(tempdir(), "climate_panel_recoded.rds"), version = 2)
```

```{r copy-file, eval=TRUE}
# not evaluated
saveRDS( climate_awareness_air, file = file.path("data-raw", "climate-panel_recoded.rds"))
```
