---
title: "Joining Harmonized Attitude Data With Air Pollution Data"
author: "Daniel Antal, CFA"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkg-load, message=FALSE}
library(dplyr)
```

In the previous blogposts we showed how to put together more than a hundred surveys conducted in about 30 countries over four years to find out about people's changing perception of the severity of climate change. We are now going to join a very different dataset, to see if we can see correlation, or even causation between people's perception of climate change and environmental degradation and their attitudes and self-reported actions.

The source of our air quality data is the European Environment Agency's (EEA) Air Quality e-Reporting (AQ e-Reporting) data on environmental pollution (particulate matter 2.5 and 10), which is recorded on the annual level for every available reporting stations and all values are in μg/m3. The period under observation is 2014-2016. 

You can get the data from European Environment Agency [here](https://www.eea.europa.eu/data-and-maps/data/aqereporting-8). _Citation: European Environment Information and Observation Network (Eionet)_

## Load the Data


```{r}
load(file.path("data", "air_pollutants.rda")) ## good practice to use system-independent file.path
climate_panel_recoded <- readRDS( file.path("data-raw", "climate-panel_recoded_agr.rds") )
names ( climate_panel_recoded)
```

```{r joining }

regional_climate_dataset <- air_pollutants %>%
  rename ( code_2016 = .data$region_nuts_codes) %>%
  left_join ( climate_panel_recoded, by = 'code_2016' ) 

summary(regional_climate_dataset)
  
```

```{r}
glm ( climate_first ~ 1- o3 + pm10 + BaP + year, data = regional_climate_dataset)

# Classification Tree with rpart
library(rpart)

# grow tree
fit <- rpart(climate_mentioned ~ o3 + pm10 + BaP + year,
   method="class", data=regional_climate_dataset)

printcp(fit) # display the results
plotcp(fit) # visualize cross-validation results
summary(fit) # detailed summary of splits

# plot tree
plot(fit, uniform=TRUE,
   main="Classification Tree for Kyphosis")
text(fit, use.n=TRUE, all=TRUE, cex=.8)

# create attractive postscript plot of tree
post(fit, file = "c:/tree.ps",
   title = "Classification Tree for Kyphosis")
```

